Run-Time Monitoring with Adjustable Overhead \\Using Dataflow-Guided Filtering

### 2015-01-05 evaluation.tex
Add discussion and eval on sub-flow dropping.

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_overhead.pdf}
    \label{fig:evaluation.umc_exec_time}
  }
  \\
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \label{fig:evaluation.umc_coverage}
  }
  \caption{Comparing dropping policies for UMC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.umc_policies}
\end{figure}

In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
The results are shown for an overhead target of 2.0x for UMC and 1.5x for BC.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overhead for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. For most benchmarks, we see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of short, independent monitoring dependence chains. Source-only dropping
causes overshoot of the overhead target for {\tt hmmer} and {\tt h264ref}.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
achieves higher coverage than unrestricted dropping while still closely matching the overhead target. For example, {\tt perlbench} shows a 10\% increase in coverage and {\tt libquantum} shows an 18\% increase in coverage by using source-only dropping. This is due
to the fact that some of the overhead of monitoring for unrestricted dropping
is being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_hb_overhead.pdf}
    \label{fig:evaluation.bc_exec_time}
  }
  \\
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_hb_coverage.pdf}
    \label{fig:evaluation.bc_coverage}
  }
  \caption{Comparing dropping policies for BC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.bc_policies}
\end{figure}

@@

One application of partial monitoring with low overhead is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overhead across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check using
unrestricted and source-only dropping policies with probabilistic dropping.
These numbers are averaged across all benchmarks. Here, we use a 10\%
probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored in at
least one of the runs. Each run
was simulated for 500 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that it is likely that at least one
monitoring event in a metadata dependence chain will be dropped with the unrestricted dropping
policy. 
Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 6\% increase in coverage with only eight runs.
With more runs, source-only dropping should be able to
reach 100\% code coverage.

>>>>

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_overhead.pdf}
    \label{fig:evaluation.umc_exec_time}
  }
  \\
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \label{fig:evaluation.umc_coverage}
  }
  \caption{Comparing dropping policies for UMC.}
  \label{fig:evaluation.umc_policies}
  \vspace{-0.2in}
  \end{center}
\end{figure}

%In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
In this section we evaluate the trade-offs between different dropping policies.
For sub-flow dropping, we identified sub-flow nodes through profiling. Because
of the need to profile the applications to determine sub-flow nodes, these
experiments were simulated for 1 billion instructions without fast-forwarding.
These results (not shown) indicated that sub-flow dropping produced similar
results in terms of coverage and overhead target matching to unrestricted
dropping for the benchmarks and monitors that we tested. Monitors and
benchmarks with difference monitoring dependence graphs may see a more
pronounced trade-off between using unrestricted dropping and sub-flow dropping.

Since, sub-flow dropping showed little difference from unrestricted dropping,
we focus on comparing unrestricted dropping and source-only dropping. The
results shown in Figures~\ref{fig:evaluation.umc_policies} and
\ref{fig:evaluation.bc_policies} were run for 2 billion instructions after
fast-forwarding for 1 billion instructions.
The results are shown for an overhead target of 2.0x for UMC and 1.5x for BC.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overhead for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. For most benchmarks, we see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of short, independent monitoring dependence chains. Source-only dropping
causes overshoot of the overhead target for {\tt hmmer} and {\tt h264ref}.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
achieves higher coverage than unrestricted dropping while still closely matching the overhead target. For example, {\tt perlbench} shows a 10\% increase in coverage and {\tt libquantum} shows an 18\% increase in coverage by using source-only dropping. This is due
to the fact that some of the overhead of monitoring for unrestricted dropping
is being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_hb_overhead.pdf}
    \label{fig:evaluation.bc_exec_time}
  }
  \\
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_hb_coverage.pdf}
    \label{fig:evaluation.bc_coverage}
  }
  \caption{Comparing dropping policies for BC.}
  \label{fig:evaluation.bc_policies}
  \vspace{-0.2in}
  \end{center}
\end{figure}

@@

One application of partial monitoring with low overhead is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overhead across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check using
unrestricted, sub-flow, and source-only dropping policies with probabilistic dropping.
These numbers are averaged across all benchmarks. Here, we use a 10\%
probability
that events are not dropped.  Each run
was simulated for 500 million instructions. 
Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored in at
least one of the runs.
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that it is likely that at least one
monitoring event in a metadata dependence chain will be dropped with the unrestricted dropping
policy. 
Instead, sub-flow dropping and source-only
dropping are much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 6\% increase in coverage with only eight runs
while sub-flow dropping shows a 1\% increase in coverage over eight runs.
While sub-flow dropping shows a slower increase in coverage than source-only
dropping, sub-flow dropping is better able to meet overhead targets.
With enough runs, both sub-flow dropping and source-only dropping should be able to
reach 100\% code coverage.

### 2015-01-05 policies.tex
Add discussion and eval on sub-flow dropping.

\begin{figure} \begin{center} \subfloat[Unrestricted Dropping]{
\includegraphics[width=\columnwidth]{figs/unrestricted_drop.pdf}
\label{fig:policies.all_drop} } \vspace{-0.1in} \\ \subfloat[Source-Only
Dropping]{ \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
\label{fig:policies.source_drop} } \vspace{-0.1in} \\ \subfloat[Sub-Flow
Dropping]{ \includegraphics[width=\columnwidth]{figs/subflow_drop.pdf}
\label{fig:policies.subflow_drop} } \end{center} \caption{Comparison of
dropping policies using metadata dependence graphs.  Square nodes represent
events where check are performed. Blue (dark) nodes indicate which nodes can be
dropped.} \label{fig:policies.policies} \end{figure}

In addition to deciding when dropping is required, trade-offs also exist in
deciding which events should be dropped.  The simplest policy is to drop any
monitoring events when slack is less than or equal to zero.  However, this can
result in wasted work. For example, consider the metadata dependence graph
shown in Figure~\ref{fig:policies.all_drop}. Here, an edge from node {\tt A} to
node {\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be dropped. Square nodes
indicate events where monitoring checks are performed. In the example, suppose
that event {\tt E} is meant to perform a check operation but is dropped.  In
this situation, the monitoring operations that were done for events {\tt C} and
{\tt D} were wasted since their results were not used for any monitoring
checks.  That is, by the time we decide to drop event {\tt E}, we have already
updated metadata for events {\tt C} and {\tt D} even though they are no longer
needed.

An alternative dropping policy which eliminates this wasted work is to only
make dropping decisions at the root or source of these metadata flows. That is,
we will decide to either monitor or not monitor an entire metadata flow. An
example of these source nodes is the monitoring done to initialize base and
bounds information on {\tt malloc} for an array bounds check. These source
nodes are typically easy to identify by the dropping hardware because they
correspond to special instructions that are used to set up metadata
information. Thus, it is not necessary to generate and analyze the monitoring
dependence graph to identify source nodes. We refer to this dropping decision
policy as \emph{source-only
dropping} and we refer to the previous policy of dropping any event as
\emph{unrestricted dropping}.

Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage at a given overhead. However, because of the coarser-grained decision, it
may be more difficult to closely match overhead targets. We expect source-only dropping 
to work well when there are a large number of independent metadata flows.
In addition, using a probabilistic dropping policy works poorly with
unrestricted dropping. Since every event in a dependent chain (e.g., events {\tt
A} through {\tt E}) needs to be monitored in order for the monitoring check to
be useful, randomly dropping any event is likely to cause the final check to be
invalid by dropping at least one event in the chain.

More complex dropping policies can be implemented given more detailed
information about the monitoring dependence graph. Here, we describe one such
policy, which we call \emph{sub-flow dropping}. Sub-flow dropping makes
dropping decisions at a granularity in between unrestricted dropping and
source-only dropping (see Figure~\ref{fig:policies.subflow_drop}). Sub-flow
dropping allows source nodes to be dropped and nodes after branch points in the
monitoring dependence graph to be dropped. From
Figure~\ref{fig:policies.subflow_drop}, this corresponds to nodes {\tt A}, {\tt
C}, and {\tt F}. The basic intuition is that dropping certain portions of the
monitoring dependence graph still result in no wasted work. For example,
dropping node {\tt C} causes nodes {\tt D} and {\tt E} to be skipped. However,
performing monitoring on the remaining nodes allows the check at node {\tt G}
to be performed with no wasted work. Similarly, dropping {\tt F} allows node
{\tt E} to be checked with no wasted work.

Depending on the target application of partial monitoring, different policies
are more applicable.
For applications where closely matching an overhead target is important, a
slack-based, unrestricted dropping policy is appropriate. However, if matching
the overhead target is not as important, then a slack-based, source-only dropping
policy could provide better coverage. 
Finally, if the goal is to use partial monitoring to enable cooperative
debugging and testing with very low overhead, then a probabilistic,
source-only dropping policy can be used to provide good total coverage over
multiple runs.
Sub-flow dropping enables a design point in between unrestricted dropping and
source-only dropping. Sub-flow dropping matches overhead better than
source-only dropping and provides better coverage than unrestricted dropping.
For a probabilistic policy, sub-flow dropping allows better overhead matching
while still allowing good total coverage over multiple runs, though at a slower
rate than source-only dropping.

>>>>

\begin{figure} 
  \begin{center} 
    \subfloat[Unrestricted Dropping] {
      \includegraphics[width=\columnwidth]{figs/unrestricted_drop.pdf}
      \label{fig:policies.all_drop} 
    }
    \vspace{-0.1in} \\ 
    \subfloat[Source-Only Dropping] { 
      \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
      \label{fig:policies.source_drop} 
    } 
    \vspace{-0.1in} \\ 
    \subfloat[Sub-Flow Dropping] { 
      \includegraphics[width=\columnwidth]{figs/subflow_drop.pdf}
      \label{fig:policies.subflow_drop} 
    } 
  \end{center} 
  \caption{Comparison of dropping policies using metadata dependence graphs.
  Square nodes represent events where check are performed. Blue (dark) nodes
  indicate which nodes can be dropped.} 
  \label{fig:policies.policies} 
\end{figure}

In addition to deciding when dropping is required, trade-offs also exist in
deciding which events should be dropped.  The simplest policy is to drop 
monitoring events when slack is less than or equal to zero.  However, this can
result in wasted work. For example, consider the metadata dependence graph
shown in Figure~\ref{fig:policies.all_drop}. Here, an edge from node {\tt A} to
node {\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be dropped. Square nodes
indicate events where monitoring checks are performed. In the example, suppose
that event {\tt E} is meant to perform a check operation but is dropped.  In
this situation, the monitoring operations that were done for events {\tt C} and
{\tt D} were wasted since their results were not used for any monitoring
checks.  That is, by the time we decide to drop event {\tt E}, we have already
updated metadata for events {\tt C} and {\tt D} even though they are no longer
needed.

An alternative dropping policy which eliminates this wasted work is to only
make dropping decisions at the root or source of these metadata flows (see Figure~\ref{fig:policies.source_drop}). That is,
we will decide to either monitor or not monitor an entire metadata flow. An
example of these source nodes is the monitoring done to initialize base and
bounds information on {\tt malloc} for an array bounds check. These source
nodes are typically easy to identify by the dropping hardware because they
correspond to the special instructions that are used to set up metadata
information. Thus, it is not necessary to generate and analyze the monitoring
dependence graph to identify source nodes. We refer to this dropping decision
policy as \emph{source-only
dropping} and we refer to the previous policy of dropping any event as
\emph{unrestricted dropping}.

More complex dropping policies can be implemented given more detailed
information about the monitoring dependence graph. Here, we describe one such
policy, which we call \emph{sub-flow dropping}. Sub-flow dropping makes
dropping decisions at a granularity in between unrestricted dropping and
source-only dropping (see Figure~\ref{fig:policies.subflow_drop}). Sub-flow
dropping allows source nodes to be dropped and nodes after branch points in the
monitoring dependence graph to be dropped. From
Figure~\ref{fig:policies.subflow_drop}, this corresponds to nodes {\tt A}, {\tt
C}, and {\tt F}. The basic idea of sub-flow dropping is to drop portions of the
monitoring dependence graph at the smallest granularity such that no work is
wasted.
For example,
dropping node {\tt C} causes nodes {\tt D} and {\tt E} to be skipped. However,
performing monitoring on the remaining nodes allows the check at node {\tt G}
to be performed with no wasted work. Similarly, dropping {\tt F} allows node
{\tt E} to be checked with no wasted work.

%Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
%comparison of where dropping decisions are made for these two policies. 
Source-only dropping will
result in no wasted work and thus better coverage at a given overhead than unrestricted dropping. However, because of the coarser-grained decision, it
may be more difficult to closely match overhead targets. 
Sub-flow dropping enables a design point in between unrestricted dropping and
source-only dropping in terms of matching overhead and coverage achieved.
Sub-flow dropping can still result in wasted work in there are merge points in
the monitoring dependence graph, but should result in less wasted work than
unrestricted dropping. Thus, its coverage falls in between that of source-only dropping and unrestricted dropping. Similarly, sub-flow dropping's ability to closely match overhead targets falls between that of source-only dropping and unrestricted dropping.
We expect source-only dropping 
to work well when there are a large number of independent metadata flows.
In addition, using a probabilistic dropping policy works poorly with
unrestricted dropping. Since every event in a dependent chain (e.g., events {\tt
A} through {\tt E}) needs to be monitored in order for the monitoring check to
be useful, randomly dropping any event is likely to cause the final check to be
invalid by dropping at least one event in the chain.

Depending on the target application of partial monitoring, different policies
are more applicable.
For applications where closely matching an overhead target is important, a
slack-based, unrestricted dropping policy is appropriate. However, if matching
the overhead target is not as important, then a slack-based, sub-flow or source-only dropping
policy could provide better coverage. 
Finally, if the goal is to use partial monitoring to enable cooperative
debugging and testing with very low overhead, then a probabilistic,
source-only dropping policy can be used to provide good total coverage over
multiple runs.
For a probabilistic policy, sub-flow dropping allows better overhead matching than source-only dropping
while still allowing good total coverage over multiple runs, though at a slower
rate than source-only dropping.
